# Inversion Method

```{R}
x=seq(-2,2,0.01)
plot(x,dnorm(x)/(pnorm(2)-pnorm(-2)),type="l")
lines(x,dunif(x,-2,2),lty=2)
# print(x)
```

```{R}
wr=function(N,M)
{
 prop=runif(N,-2,2)
 weight=dnorm(prop) #I’ve cheated by using the intrinsic dnorm command!
 x=sample(prop,M,TRUE,weight)  #Ok to supply the unormalised weights
 return(x)
}
#example
hist(wr(50,50),freq=F)
lines(x,dnorm(x)/(pnorm(2)-pnorm(-2)),type="l")
```

# Accept/Rejection algorithm

Suppose the objective function is:

$$
f(x) = \frac{6(x-0.5)^2}{7},\quad x \in (0,2)
$$

The proposal distribution is:

$$
g(x) = 0.5
$$

Since that $\max_{0\leq x \leq 2} f(x) = 1.9286$, we set $M = 3.858$

```{r}
N <- 50000   # set the number of random numbers to generate to 500,000
M <- 3.858    # set the value of M used in the calculation to 3.858
y <- runif(N, min = 0, max = 2)  # generate N uniformly distributed random numbers between 0 and 2, and assign them to the variable y
u <- runif(N, min = 0, max = 1)  # generate N uniformly distributed random numbers between 0 and 1, and assign them to the variable u
gy <- 0.5    # set the value of gy used in the calculation to 0.5
fy <- 6 * (y - 0.5)^2 / 7   # calculate the value of fy for each element of y using the given formula
x <- y[u < fy / gy / M]   # generate a random sample of numbers from y where u is less than fy/(gy*M)
sample <- length(x)    # calculate the length of the generated sample and assign it to the variable sample
hist(x, breaks = 50, freq = FALSE, col = "#adabab",
     main = "f(x) = 6(x - 0.5)^2/7")   # plot a histogram of the generated sample using 50 bins, with a grey color and no frequency labels, and set the title of the plot to "f(x) = 6(x - 0.5)^2/7"

curve(6 * (x - 0.5)^2 / 7, from = 0, to = 2, add = TRUE, col = "red")  # add a curve to the existing plot showing the true PDF with a red color

```

The code generates a random sample of numbers from a PDF using the acceptance-rejection method. The PDF is defined as f(x) = 6(x - 0.5)^(2/7) in the code. The code then plots a histogram of the generated sample along with the true PDF. The histogram is plotted using the hist function, and the true PDF is plotted using the curve function. The add = TRUE argument is used to add the curve to the existing plot, rather than creating a new plot. The col argument is used to set the color of the histogram and curve.

## MCMC

```{r}
f <- c(0, 1, 2, 3, 4, 5, 6, 5, 4, 3, 2, 1)
N <- 100000
d <- rep(0, N)
x <- 5
for (i in 1:N) {
  U <- runif(1)
  if (x == 2) {
    if (U < 0.5) {
      y <- 3
    } else {
      y <- 2
    }
  } else if (x == 12) {
    if (U < 0.5) {
      y <- 11
    } else {
      y <- 12
    }
  } else {
    if (U < 0.5) {
      y <- x - 1
    } else {
      y <- x + 1
    }
  }
  h <- min(1, f[y] / f[x])
  U <- runif(1)
  if (U < h) {
    x <- y
  }
  d[i] <- x
}
a <- 1:12
hist(d, breaks = a)

```